---
title: "Graphs"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(party)
library(ggridges)
library(factoextra)
library(knitr)
library(anchors)
library(xtable)
Sys.setlocale("LC_CTYPE", "russian")

si <- read.csv('si.csv')
sk <- read.csv('sk.csv')
gi <- read.csv('gi.csv')
sa <- read.csv('sa.csv')

sa$dial <- as.factor(sa$dia)
si$dial <- as.factor(si$dia)
sk$dial <- as.factor(sk$dia)
gi$dial <- as.factor(gi$dia)

sa$norm <- as.factor(sa$norm)
si$norm <- as.factor(si$norm)
sk$norm <- as.factor(sk$norm)
gi$norm <- as.factor(gi$norm)


```


# Reflection of *s in anlaut

## Table
Dialects: v, d, i, p, s, t
```{r warning = FALSE, echo = FALSE, results='asis'}
kable(table(sa$dial, sa$norm))
```

```{r warning = F, echo = F}
sa %>% ggplot(aes(x = dial, fill = norm)) + geom_bar() +
  labs(x = 'dialect') +
  scale_y_continuous(breaks = seq(0, 80, 5)) +
  geom_text(aes(label=..count..),stat="count",position=position_stack())
```

## Overall density plots
```{r warning = F, echo = F}
sa %>% ggplot(aes(dur)) +
  geom_density(fill = 'blue', alpha = 0.5) +
  scale_x_continuous(breaks = seq(0, 0.6, 0.03)) +
  labs(x = 'Duration')
```

```{r warning = F, echo = F}
sa %>% ggplot(aes(int)) +
  geom_density(fill = 'blue', alpha = 0.5) +
  scale_x_continuous(breaks = seq(20, 70, 2)) +
  labs(x = 'Intensity')
```



## Statistics
Dialects: v, d, i, p, s, t
```{r warning = F}
summary(sa$int)
summary(sa$dur)
sa %>% group_by(dial) %>% 
  summarise(mean_dur = mean(dur),
            sd_dur = sd(dur),
            mean_int = mean(int),
            sd_int = sd(int)) %>% 
  mutate(across(where(is.numeric), round, 3))
cor.test(sa$int, sa$dur)

```

## Facet tables
```{r warning = F, echo = F, fig.width = 10, fig.height=8}
sa %>% ggplot(aes(dur)) +
  geom_histogram(fill = 'white', color = 'black', bins = 30) +
  facet_wrap(~dial) + labs(x = 'Duration') +
  scale_x_continuous(breaks = seq(0, 0.6, 0.05)) +
  theme(axis.text.x = element_text(angle = 90))
```

```{r warning = F, echo = F}
sa %>% ggplot(aes(int)) +
  geom_histogram(fill = 'white', color = 'black', bins = 30) +
  facet_wrap(~dial) + labs(x = 'Intensity') + 
  scale_x_continuous(breaks = seq(20, 70, 5))
```


## Boxplots
```{r warning = F, echo=F}
sa %>% ggplot(aes(y = dur, fill = dial, x = dial)) +
  geom_boxplot() + labs(y = 'Duration', x = 'dialect',
                        fill = 'dialect') +
  theme(axis.title.x=element_blank()) + 
  scale_y_continuous(breaks = seq(0, 0.6, 0.05))
```

```{r warning = F, echo=F}
sa %>% ggplot(aes(y = int, fill = dial, x = dial)) +
  geom_boxplot() + labs(y = 'Intensity', x = 'dialect',
                        fill = 'dialect') +
  theme(axis.title.x=element_blank()) + 
  scale_y_continuous(breaks = seq(20, 70, 5))
```



## Scatter plots
```{r warning=F, echo=F}
sa %>% 
  ggplot(aes(dur, int)) +
  geom_point(aes(color = norm)) + geom_smooth(alpha=0.2)+
labs(color = 'Norm', x = 'Duration', y = 'Intensity') +
  scale_y_continuous(breaks = seq(20, 70, 5)) + 
  scale_x_continuous(breaks = seq(0, 0.6, 0.05))
```

```{r warning=F, echo=F}
sa %>% 
  ggplot(aes(dur, int)) +
  geom_point(aes(color = dial)) + geom_smooth(alpha=0.2) +
  labs(color = 'dialect', x = 'Duration', y = 'Intensity') +
  scale_y_continuous(breaks = seq(20, 70, 5)) + 
  scale_x_continuous(breaks = seq(0, 0.6, 0.05))
```

## Tree decision
```{r warning = F, echo = F}
fit <- ctree(dial~dur + int, data = sa)
plot(fit)
```

# Reflection of *s in intervocalic inlaut

## Table
```{r warning = FALSE, echo = FALSE, results='asis'}
kable(table(si$dia, si$norm))
```

```{r warning = F, echo = F}
si %>% ggplot(aes(x = dial, fill = norm)) + geom_bar() +
  labs(x = 'dialect') +
  scale_y_continuous(breaks = seq(0, 80, 5)) +
  geom_text(aes(label=..count..),stat="count",position=position_stack())
```

## Overall density plots
```{r warning = F, echo = F}
si %>% ggplot(aes(dur)) +
  geom_density(fill = 'blue', alpha = 0.5) +
  scale_x_continuous(breaks = seq(0, 0.6, 0.03)) +
  labs(x = 'Duration')
```

```{r warning = F, echo = F}
si %>% ggplot(aes(int)) +
  geom_density(fill = 'blue', alpha = 0.5) +
  scale_x_continuous(breaks = seq(20, 70, 2)) +
  labs(x = 'Intensity')
```



## Statistics
```{r warning = F}
summary(si$int)
summary(si$dur)
kable(si %>% group_by(dial) %>% 
  summarise(mean_dur = mean(dur),
            sd_dur = sd(dur),
            mean_int = mean(int),
            sd_int = sd(int)) %>% 
  mutate(across(where(is.numeric), round, 3)))
cor.test(si$int, si$dur)

```

## Facet tables
```{r warning = F, echo = F, fig.width = 10, fig.height=8}
si %>% ggplot(aes(dur)) +
  geom_histogram(fill = 'white', color = 'black', bins = 30) +
  facet_wrap(~dial) + labs(x = 'Duration') +
  scale_x_continuous(breaks = seq(0, 0.6, 0.05)) +
  theme(axis.text.x = element_text(angle = 90))
```

```{r warning = F, echo = F}
si %>% ggplot(aes(int)) +
  geom_histogram(fill = 'white', color = 'black', bins = 30) +
  facet_wrap(~dial) + labs(x = 'Intensity') + 
  scale_x_continuous(breaks = seq(20, 70, 5))
```


## Boxplots
```{r warning = F, echo=F}
si %>% ggplot(aes(y = dur, fill = dial, x = dial)) +
  geom_boxplot() + labs(y = 'Duration', x = 'dialect',
                        fill = 'dialect') +
  theme(axis.title.x=element_blank()) + 
  scale_y_continuous(breaks = seq(0, 0.6, 0.05))
```

```{r warning = F, echo=F}
si %>% ggplot(aes(y = int, fill = dial, x = dial)) +
  geom_boxplot() + labs(y = 'Intensity', x = 'dialect',
                        fill = 'dialect') +
  theme(axis.title.x=element_blank()) + 
  scale_y_continuous(breaks = seq(20, 70, 5))
```



## Scatter plots
```{r warning=F, echo=F}
si %>% 
  ggplot(aes(dur, int)) +
  geom_point(aes(color = norm)) + geom_smooth(alpha=0.2)+
labs(color = 'Norm', x = 'Duration', y = 'Intensity') +
  scale_y_continuous(breaks = seq(20, 70, 5)) + 
  scale_x_continuous(breaks = seq(0, 0.6, 0.05))
```

```{r warning=F, echo=F}
si %>% 
  ggplot(aes(dur, int)) +
  geom_point(aes(color = dial)) + geom_smooth(alpha=0.2) +
  labs(color = 'dialect', x = 'Duration', y = 'Intensity') +
  scale_y_continuous(breaks = seq(20, 70, 5)) + 
  scale_x_continuous(breaks = seq(0, 0.6, 0.05))
```

<!-- ## Tree decision -->
```{r warning = F, echo = F}
# fit <- ctree(dial~dur + int, data = si)
# plot(fit)
```


# Reflection of *s in consonant cluster -sk-/-ks-

## Table
```{r warning = FALSE, echo = FALSE, results='asis'}
kable(table(sk$dia, sk$norm))
```

```{r warning = F, echo = F}
sk %>% ggplot(aes(x = dial, fill = norm)) + geom_bar() +
  labs(x = 'dialect') +
  scale_y_continuous(breaks = seq(0, 80, 5)) +
  geom_text(aes(label=..count..),stat="count",position=position_stack())
```

## Overall density plots
```{r warning = F, echo = F}
sk %>% ggplot(aes(dur)) +
  geom_density(fill = 'blue', alpha = 0.5) +
  scale_x_continuous(breaks = seq(0, 0.6, 0.02)) +
  labs(x = 'Duration')
```

```{r warning = F, echo = F}
sk %>% ggplot(aes(int)) +
  geom_density(fill = 'blue', alpha = 0.5) +
  scale_x_continuous(breaks = seq(20, 75, 2)) +
  labs(x = 'Intensity')
```



## Statistics
```{r warning = F}
summary(sk$int)
summary(sk$dur)
kable(sk %>% group_by(dial) %>% 
  summarise(mean_dur = mean(dur),
            sd_dur = sd(dur),
            mean_int = mean(int),
            sd_int = sd(int)) %>% 
  mutate(across(where(is.numeric), round, 3)))
cor.test(sk$int, sk$dur)

```

## Facet tables
```{r warning = F, echo = F, fig.width = 10, fig.height=8}
sk %>% ggplot(aes(dur)) +
  geom_histogram(fill = 'white', color = 'black', bins = 30) +
  facet_wrap(~dial) + labs(x = 'Duration') +
  scale_x_continuous(breaks = seq(0, 0.6, 0.02)) +
  theme(axis.text.x = element_text(angle = 90))
```

```{r warning = F, echo = F}
sk %>% ggplot(aes(int)) +
  geom_histogram(fill = 'white', color = 'black', bins = 30) +
  facet_wrap(~dial) + labs(x = 'Intensity') + 
  scale_x_continuous(breaks = seq(20, 73, 5))
```


## Boxplots
```{r warning = F, echo=F}
sk %>% ggplot(aes(y = dur, fill = dial, x = dial)) +
  geom_boxplot() + labs(y = 'Duration', x = 'dialect',
                        fill = 'dialect') +
  theme(axis.title.x=element_blank()) + 
  scale_y_continuous(breaks = seq(0, 0.6, 0.02))
```

```{r warning = F, echo=F}
sk %>% ggplot(aes(y = int, fill = dial, x = dial)) +
  geom_boxplot() + labs(y = 'Intensity', x = 'dialect',
                        fill = 'dialect') +
  theme(axis.title.x=element_blank()) + 
  scale_y_continuous(breaks = seq(20, 75, 5))
```



## Scatter plots
```{r warning=F, echo=F}
sk %>% 
  ggplot(aes(dur, int, color = norm)) +
  geom_point(aes(color = norm)) + geom_smooth(alpha=0.2, method = 'lm')+
labs(color = 'Norm', x = 'Duration', y = 'Intensity') +
  scale_y_continuous(breaks = seq(20, 70, 5)) + 
  scale_x_continuous(breaks = seq(0, 0.6, 0.05))
```

```{r warning=F, echo=F}
sk %>% 
  ggplot(aes(dur, int)) +
  geom_point(aes(color = dial)) + geom_smooth(alpha=0.2) +
  labs(color = 'dialect', x = 'Duration', y = 'Intensity') +
  scale_y_continuous(breaks = seq(20, 70, 5)) + 
  scale_x_continuous(breaks = seq(0, 0.6, 0.02))
```

## Tree decision
```{r warning = F, echo = F}
fit <- ctree(dial~dur + int, data = sk)
plot(fit)
```

# Reflection of *g in intervocalic inlaut

## Table
```{r warning = FALSE, echo = FALSE, results='asis'}
kable(table(gi$dia, gi$norm))
```

```{r warning = F, echo = F}
gi %>% ggplot(aes(x = dial, fill = norm)) + geom_bar() +
  labs(x = 'dialect') +
  scale_y_continuous(breaks = seq(0, 80, 5)) +
  geom_text(aes(label=..count..),stat="count",position=position_stack())
```

## Overall density plots
```{r warning = F, echo = F}
gi %>% ggplot(aes(dur)) +
  geom_density(fill = 'blue', alpha = 0.5) +
  scale_x_continuous(breaks = seq(0, 0.6, 0.02)) +
  labs(x = 'Duration')
```

```{r warning = F, echo = F}
gi %>% ggplot(aes(int)) +
  geom_density(fill = 'blue', alpha = 0.5) +
  scale_x_continuous(breaks = seq(20, 120, 2)) +
  labs(x = 'Intensity')
```



## Statistics
```{r warning = F}
summary(gi$int)
summary(gi$dur)
kable(gi %>% group_by(dial) %>% 
  summarise(mean_dur = mean(dur),
            sd_dur = sd(dur),
            mean_int = mean(int),
            sd_int = sd(int)) %>% 
  mutate(across(where(is.numeric), round, 3)))
cor.test(gi$int, gi$dur)

```

## Facet tables
```{r warning = F, echo = F, fig.width = 10, fig.height=8}
gi %>% na.omit() %>%  ggplot(aes(dur)) +
  geom_histogram(fill = 'white', color = 'black', bins = 30) +
  facet_wrap(~dial) + labs(x = 'Duration') +
  scale_x_continuous(breaks = seq(0, 0.6, 0.02)) +
  theme(axis.text.x = element_text(angle = 90))
```

```{r warning = F, echo = F}
gi %>% na.omit() %>%  ggplot(aes(int)) +
  geom_histogram(fill = 'white', color = 'black', bins = 30) +
  facet_wrap(~dial) + labs(x = 'Intensity') + 
  scale_x_continuous(breaks = seq(20, 90, 5))
```


## Boxplots
```{r warning = F, echo=F}
gi %>% na.omit() %>% 
  ggplot(aes(y = dur, fill = dial, x = dial)) +
  geom_boxplot() + labs(y = 'Duration', x = 'dialect',
                        fill = 'dialect') +
  theme(axis.title.x=element_blank()) + 
  scale_y_continuous(breaks = seq(0, 0.6, 0.02))
```

```{r warning = F, echo=F}
gi %>% na.omit() %>% 
  ggplot(aes(y = int, fill = dial, x = dial)) +
  geom_boxplot() + labs(y = 'Intensity', x = 'dialect',
                        fill = 'dialect') +
  theme(axis.title.x=element_blank()) + 
  scale_y_continuous(breaks = seq(20, 90, 5))
```



## Scatter plots
```{r warning=F, echo=F}
gi %>% 
  ggplot(aes(dur, int)) +
  geom_point(aes(color = norm)) + geom_smooth(alpha=0.2)+
labs(color = 'Norm', x = 'Duration', y = 'Intensity') +
  scale_y_continuous(breaks = seq(20, 90, 5)) + 
  scale_x_continuous(breaks = seq(0, 0.6, 0.02))
```

```{r warning=F, echo=F}
gi %>% na.omit() %>% 
  ggplot(aes(dur, int)) +
  geom_point(aes(color = dial)) + geom_smooth(alpha=0.2) +
  labs(color = 'dialect', x = 'Duration', y = 'Intensity') +
  scale_y_continuous(breaks = seq(20, 90, 5)) + 
  scale_x_continuous(breaks = seq(0, 0.6, 0.02))
```

## Tree decision
```{r warning = F, echo = F}
fit <- ctree(dial~dur + int, data = na.omit(gi))
plot(fit)
```


## GI & SI

Negative correlation between both gi and si
```{r}
cor.test(gi$dur, gi$int)
cor.test(si$dur, si$int)
```

## Intensity and Duration interaction between groups
```{r warning=F, echo=F}
rbind(si %>% mutate(group = 's'), 
      gi %>% mutate(group = 'g')) %>% 
  ggplot(aes(dur, int, color = group)) +
  geom_point() + 
  geom_smooth(se = F) +
  scale_y_continuous(breaks = seq(20, 90, 5)) + 
  scale_x_continuous(breaks = seq(0, 0.6, 0.02)) +
  labs(x = 'Duration', y = 'Intensity')
```


```{r echo = FALSE, warning = FALSE}
rbind(si %>% mutate(group = 's'), 
      gi %>% mutate(group = 'g')) %>% 
  ggplot(aes(dur, int, color = group)) +
  geom_point() + 
  geom_smooth(se = F, method = 'lm')+
  scale_y_continuous(breaks = seq(20, 90, 5)) + 
  scale_x_continuous(breaks = seq(0, 0.6, 0.02))
```

## Boxplots
```{r warning=F, echo=F}
rbind(si %>% mutate(group = 's'), 
      gi %>% mutate(group = 'g')) %>% 
  na.omit() %>% 
  ggplot(aes(dial, dur, fill = group)) +
  geom_boxplot() +
  scale_y_continuous(breaks = seq(0, 0.6, 0.025)) +
  labs(y = 'Duration')
```


```{r warning = F, echo=F}
rbind(si %>% mutate(group = 's'), 
      gi %>% mutate(group = 'g')) %>% 
  na.omit() %>% 
  ggplot(aes(dial, int, fill = group)) +
  geom_boxplot() +
  scale_y_continuous(breaks = seq(0, 100, 5)) +
  labs(y = 'Intensity')
```

